Probabilistic Model of Bilateral Lymphatic Spread in Head and Neck Cancer

### 2024-04-15 manuscript.qmd
continue with data section/table

data = pd.read_csv(paths.data, header=[0,1,2])
data[COL.t_stage] = data[COL.t_stage].apply(lambda x: "early" if x <= 2 else "late")
early_midext_data = (data.loc[data[COL.t_stage] == "early", COL.midext] == True).mean()
late_midext_data = (data.loc[data[COL.t_stage] == "late", COL.midext] == True).mean()
grouped = data.groupby(by=[COL.midext, COL.t_stage, COL.ipsi_III])

@@

The data used to train our models consists of <!-- `{python} len(data)` --> patients with newly diagnosed oropharyngeal squamous cell carcinomas. It reports the involvement of every patient individually and per lymph node level in tabular form, in addition to other clinico-pathological information such as age, T-category, and HPV p16 status. Their patient records have been collected at four different instituions:

1. <!-- `{python} len(usz_data)` --> patients from the University Hospital Zurich [@ludwig_dataset_2022].
2.

>>>>

raw = pd.read_csv(paths.data, header=[0,1,2])
data = raw.copy()
data[COL.t_stage] = data[COL.t_stage].apply(lambda x: "early" if x <= 2 else "late")
early_midext_data = (data.loc[data[COL.t_stage] == "early", COL.midext] == True).mean()
late_midext_data = (data.loc[data[COL.t_stage] == "late", COL.midext] == True).mean()
grouped = data.groupby(by=[COL.midext, COL.t_stage, COL.ipsi_III])

@@

The data used to train our models consists of <!-- {python} len(data) --> patients with newly diagnosed oropharyngeal squamous cell carcinomas. It reports the involvement of every patient individually and per lymph node level in tabular form, in addition to other clinico-pathological information such as age, T-category, and HPV p16 status. Their patient records have been collected at four different instituions:

```{python}
#| echo: false
#| label: tbl-data

col_map = {
  ("patient", "#", "institution"): ("Institution", lambda x: None),
  ("patient", "#", "age"): ("Age", np.mean),
  ("tumor", "1", "t_stage"): ("T-category", np.mean),
}
reduced = raw[col_map.keys()]
reduced.columns = reduced.columns.get_level_values(2)
renamed = (
  reduced
  .rename(columns={tpl[-1]: val[0] for tpl, val in col_map.items()})
  .convert_dtypes()
)
grouped = renamed.groupby(by="Institution")
aggregated = grouped.agg({
  name: func for name, func in col_map.values()
  if name != "Institution"
})
```

```{python}
#| echo: false
#| output: asis

for i, institution in enumerate([
  "University Hospital Zurich",
  "Centre Léon Bérard",
  "Inselspital Bern",
  "Vall d'Hebron Barcelona Hospital",
]):
  is_from_institution = data["patient", "#", "institution"] == institution
  inst_data = data.loc[is_from_institution]
  locations = inst_data["tumor", "1", "location"].unique()
  locations = locations[~pd.isna(locations)]

  if len(locations) == 1:
    locations_str = str(locations[0])
  else:
    locations_str = ", ".join(locations[:-1]) + ", and " + str(locations[-1])

  print(f"{i}. {len(inst_data)} patients with newly diagnosed HNSCC in the {locations_str}, extracted at the {institution}.")
```
